---
title: "Differential expression"
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

```{r loadLibraries, echo=FALSE, warning=FALSE, message=FALSE}
library(rmarkdown)
library(tidyverse)
library(stringr)
library(DESeq2)
library(tinytex)
library(knitr)
library(ggplot2)
library(plotly)
library(pheatmap)
library(RColorBrewer)
library(readr)
library(DT)
library(pcaExplorer)
library(ashr)
library(EnhancedVolcano)
```

[//]: # "Load the various chunks in the various files"
```{r readChunks, echo=FALSE, warning=FALSE, message=FALSE}
read_chunk( path=file.path( SCRIPT_DIR, "data_quality.R"))
read_chunk( path=file.path( SCRIPT_DIR, "diffexp.R"))
```
# Data quality assessment by sample clustering and visualization

Here we analyze the similarities and variabilities between the different conditions.

We regularized log transformation with rlog.
This function transforms the count data to the log2 scale in a way which minimizes
differences between samples for rows with small counts, and which normalizes with 
respect to library size. The rlog transformation is more robust in the case
when the size factors vary widely. The transformation is useful when checking for outliers 

## Heatmap of the count matrix

- Heatmap of the 30 genes with the top and bottom loadings.
(i.e. the genes that explain the best the variability on the PCA component).

```{r heatmap-count, echo=FALSE, results = "asis", message=FALSE}
<<heatmap-count>>
```

## Principal component plot of the samples

We report for each 2 first components of the PCA the 30 genes with the  top and bottom loadings.

```{r plot-pca, echo=FALSE, results = "asis"}
<<heatmap-count>>
<<plot-pca>>
<<heatmap-distance>>
```

## Heatmap of the sample-to-sample distances

We apply the dist function for the transpose of the transformed count matrix to get sample-to-sample distances.
A heatmap of this distance matrix gives us an overview over similarities and dissimilarities between samples. 

```{r heatmap-distance, echo=FALSE, results = "asis"}
<<heatmap-distance>>
```

# Differential expression analysis

In this section we execute an analysis of the differentially expressed genes 
between the different conditions.

## MA plot and Volcano plot

MA plot representing the log2 fold changes attributable to a given variable over
the mean of normalized counts for each conditions. 
Points will be colored in blue if the adjusted p value is less than 0.1.
Points which fall out of the window are plotted as open triangles pointing 
either up or down. All genes are presented in the MA plots

Volcanoplot representing the significance versus fold-change on the y and x axes.

```{r diffexp, echo=FALSE, results = "asis", message=FALSE, warning = FALSE}
<<diffexp>>
```

